Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-09 ICPP-19/communications.tex
minor
THe optimal schedule  does not have same order on both resources

@@

		\begin{figure}[htb]
			\tikzset{xtick/.style={inner xsep=0pt, inner ysep=3pt, minimum size=0pt, draw},%
				task/.style args={#1start#2length#3res#4color#5}{rounded corners, draw, inner
					sep=0pt, fill=#5, label=center:#1, fit={(#2,#4*0.75) (#2+#3,#4*0.75+0.75)}},%
				vert/.style={inner sep=1pt, fill=black, circle, draw, label=#1}
			}
			\newcommand{\schedule}[1]{
				\draw[->] (-0.4, 0) -- (#1, 0) node[below] {$t$};
				\draw (0, 0) -- (0, 1.5) node[pos=0.25, left] {Comp.}
				node[pos=0.75, left] {Comm.};
				\draw[dashed,gray] (0, 0.75) -- (#1, 0.75);
			}
			\centering
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.6]
			\schedule{12.5}
			\node[task=$A_{1,1}$ start 0 length 1 res 1 color cyan]{};
			\node[task=$A_{1,2}$ start 1 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 2 length 1 res 1 color blue!70!white]{};
			\node[task=$K_0$ start 0 length 3 res 0 color gray!40!white]{};
			\node[task=$K_1$ start 3 length 6 res 1 color green]{}; 
			\node[task=$A_{1,1}$ start 3 length 1.8 res 0 color cyan]{};
			\node[task=$A_{1,2}$ start 4.8 length 2.3 res 0 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 7.1 length 1.9 res 0 color blue!70!white]{};
			\node[task=$A_{2,1}$ start 9 length 1 res 1 color cyan]{};
			\node[task=$A_{2,2}$ start 10 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{2,3}$ start 11 length 1 res 1 color blue!70!white]{};
			\node[task=$K_1$ start 9 length 3 res 0 color green]{};
			\draw[<->,thin] (0, -0.2) -- node[below]{$3$} (3, -0.2) ;
			\draw[<->,thin] (9, -0.2) -- node[below]{$3$} (12, -0.2) ;
			\draw[<->,thin] (3, -0.2) -- node[below]{$b'$} (9, -0.2) ;
			\end{tikzpicture}
			\caption{ \label{fig:firstSegment} Pattern of feasible schedule $S$.}
		\end{figure}

@@

@@

	\tikzset{xtick/.style={inner xsep=0pt, inner ysep=3pt, minimum
			size=0pt, draw, label=below:#1},%
		comm/.style args={#1start#2length#3color#4}{rounded corners=1mm, draw, inner
			sep=0pt, fill=#4, label=center:#1, fit={(#2,0.75)
				(#2+#3,1.5)}},%
		comp/.style args={#1start#2length#3color#4}{rounded corners=1mm, draw, inner
			sep=0pt, fill=#4, label=center:#1, fit={(#2,0)
				(#2+#3,0.75)}},%
	}
	\newcommand{\schedule}[3]{
		\draw[->] (-0.2, 0) -- (#1, 0) node[below] {$t$};
		\draw (0, 0) -- (0, 1.5);
		\node at (-0.8, 0.75)[rotate=90] {#2};
		\draw[dashed,gray] (0, 0.75) -- (#1, 0.75);
		\foreach \t in {0,#3} {
			\node[xtick=\t] at (\t, 0){};
		}
	}
	\newcommand{\task}[6][0]{
		\node[comm=#2 start #3 length #4 color #6]{};
		\node[comp=#2 start #3+#4+#1 length #5 color #6]{}; 
	}

@@

		\todo[inline]{increase font size in all plots, may be free scale-y}
		%%	\todo[inline]{SKumar: decide metrics: ratio to optimal, percentage overlap, 100 - percentage non overlap}

@@

		\subsection{Scheduling in Batches}
		In most applications, scheduler may only observe a limited batch of independent tasks. Therefore we organize tasks of each trace file in the batch of 100. Last group may have less than 100 tasks. We apply each heuristic on each group consequently. Figure~\ref{fig:best_variants_batch} shows performance of the best variants of all categories for both applications. The plots exhibit behavior similar to Figures~\ref{fig:ratio_to_optimal_best_hf} and ~\ref{fig:ratio_to_optimal_best_ccsd}, static order with dynamic correction variants attain maximum communication-computation overlap and outperform other heuristics.
		\todo[inline]{SKumar:combine both plots}

>>>>

% Command for figures

\tikzset{xtick/.style={inner xsep=0pt, inner ysep=3pt, minimum
		size=0pt, draw, label=below:#1},%
	comm/.style args={#1start#2length#3color#4}{rounded corners=1mm, draw, inner
		sep=0pt, fill=#4, label=center:#1, fit={(#2,0.75)
			(#2+#3,1.5)}},%
	comp/.style args={#1start#2length#3color#4}{rounded corners=1mm, draw, inner
		sep=0pt, fill=#4, label=center:#1, fit={(#2,0)
			(#2+#3,0.75)}},%
}

\newcommand{\schedule}[3]{
	\draw[->] (-0.2, 0) -- (#1, 0) node[below] {$t$};
	\draw (0, 0) -- (0, 1.5);
	\node at (-0.8, 0.75)[rotate=90] {#2};
	\draw[dashed,gray] (0, 0.75) -- (#1, 0.75);
	\foreach \t in {0,#3} {
		\node[xtick=\t] at (\t, 0){};
	}
}
\newcommand{\task}[6][0]{
	\node[comm=#2 start #3 length #4 color #6]{};
	\node[comp=#2 start #3+#4+#1 length #5 color #6]{}; 
}

@@

		\begin{figure}[htb]
			\tikzset{xtick/.style={inner xsep=0pt, inner ysep=3pt, minimum size=0pt, draw},%
				task/.style args={#1start#2length#3res#4color#5}{rounded corners, draw, inner
					sep=0pt, fill=#5, label=center:#1, fit={(#2,#4*0.75) (#2+#3,#4*0.75+0.75)}},%
				vert/.style={inner sep=1pt, fill=black, circle, draw, label=#1}
			}
			\newcommand{\scheduleNoName}[1]{
				\draw[->] (-0.4, 0) -- (#1, 0) node[below] {$t$};
				\draw (0, 0) -- (0, 1.5) node[pos=0.25, left] {Comp.}
				node[pos=0.75, left] {Comm.};
				\draw[dashed,gray] (0, 0.75) -- (#1, 0.75);
			}
			\centering
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.6]
			\scheduleNoName{12.5}
			\node[task=$A_{1,1}$ start 0 length 1 res 1 color cyan]{};
			\node[task=$A_{1,2}$ start 1 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 2 length 1 res 1 color blue!70!white]{};
			\node[task=$K_0$ start 0 length 3 res 0 color gray!40!white]{};
			\node[task=$K_1$ start 3 length 6 res 1 color green]{}; 
			\node[task=$A_{1,1}$ start 3 length 1.8 res 0 color cyan]{};
			\node[task=$A_{1,2}$ start 4.8 length 2.3 res 0 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 7.1 length 1.9 res 0 color blue!70!white]{};
			\node[task=$A_{2,1}$ start 9 length 1 res 1 color cyan]{};
			\node[task=$A_{2,2}$ start 10 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{2,3}$ start 11 length 1 res 1 color blue!70!white]{};
			\node[task=$K_1$ start 9 length 3 res 0 color green]{};
			\draw[<->,thin] (0, -0.2) -- node[below]{$3$} (3, -0.2) ;
			\draw[<->,thin] (9, -0.2) -- node[below]{$3$} (12, -0.2) ;
			\draw[<->,thin] (3, -0.2) -- node[below]{$b'$} (9, -0.2) ;
			\end{tikzpicture}
			\caption{ \label{fig:firstSegment} Pattern of feasible schedule $S$.}
		\end{figure}

@@

	This Theorem shows that adding a memory constraint to our
	problem makes it more difficult. One additional difficulty
	compared to limited capacity 2-machine
	flowshop~\cite{Papadimitriou:1980:FSL:322203.322213} is that
	it may not be optimal to consider the same ordering on both
	machines:

	\begin{proposition}
		There exists an instance of $DT$ for which in all optimal
		schedules, the communication order of tasks is different
		from their computation order.
	\end{proposition}

	\begin{table}
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				%%				\multirow{2}{*}{Task} & Comm Time & \multirow{2}{*}{Comp Time}\\ 
				%%				&=Comm Volume& \\ \hline
				Task & Comm Volume & Comm Time & Comp Time\\ \hline 
				A & 0 & 0 & 5\\ \hline
				B & 4 & 4 & 3\\ \hline
				C & 1 & 1 & 6\\ \hline
				D & 3 & 3 & 7\\ \hline
				E & 6 & 6 & 0.5\\ \hline
				F & 7 & 7 & 0.5\\ \hline
			\end{tabular}
		\end{center}
		\caption{Example instance where ordering on both resources
			has to be different.}
		\label{table:different.order}
	\end{table}

	\begin{figure}
		\centering
		\newcommand{\taskA}[1]{\node[comp=$A$ start #1 length 5 color cyan] {};}
		\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{4}{3}{blue!25!white}}
		\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{1}{6}{blue!50!white}}
		\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{3}{7}{blue!75!white}}
		\newcommand{\taskE}[2][0]{\node[comm=$E$ start #2 length 6 color green!25!white]{};%
			\node[comp={} start #2+6+#1 length 0.5 color green!25!white]{};}
		\newcommand{\taskF}[2][0]{\node[comm=$F$ start #2 length 7 color green!50!white]{};%
			\node[comp={} start #2+7+#1 length 0.5 color green!50!white]{};}

		\subfloat[][Optimal schedule with common ordering on both
		resources]{
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.3]
			\schedule{24}{}{5,8,15,21.5,23}
			\taskA{0}
			\taskB[1]{0}
			\taskD[1]{4}
			\taskE[1]{8}
			\taskC[0.5]{14}
			\taskF{15.5}
			\end{tikzpicture}
			\label{fig:different.order.same}
		}

		\subfloat[][Schedule with different ordering on both
		resources]{
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.3]
			\schedule{24}{}{5,8,14,22}
			\taskA{0}
			\taskB[1]{0}
			\taskC[3]{4}
			\taskD[6.5]{5}
			\taskE{8}
			\taskF{14.5}
			\end{tikzpicture}
			\label{fig:different.order.opt}
		}
		\caption{Schedules for the instance of
			Table~\ref{table:different.order}}
		\label{fig:different.order}
	\end{figure}

	\begin{proof}
		Consider the instance described on
		Table~\ref{table:different.order}, in which memory capacity
		is $C=10$. Figure~\ref{fig:different.order.same} shows the
		best possible schedule when tasks are scheduled in the same
		order on both resources (obtained by exhaustive search). On
		the other hand, Figure~\ref{fig:different.order.opt} shows
		another schedule with lower makespan, in which the order is
		different.

		In the infinite memory case, the standard proof that an
		optimal schedule exists with the same order on both
		resources claims that it is possible to swap two tasks which
		do not satisfy this property. On
		Figure~\ref{fig:different.order}, this would mean swapping
		tasks $D$ and $E$. But the communication of task $E$ can not
		start earlier because it would not fit in memory with tasks
		$B$ and $C$, and delaying the computation of task $E$ after
		task $D$ would delay task $F$ because $E$ and $F$ do not fit
		in memory together. We can see that this claim does not hold
		in the constrained memory case. 
	\end{proof}

@@

@@

		%%		\todo[inline]{increase font size in all plots, may be free scale-y}
		%%	\todo[inline]{SKumar: decide metrics: ratio to optimal, percentage overlap, 100 - percentage non overlap}

@@

		\subsection{Scheduling in Batches}
		In most applications, scheduler may only observe a limited batch of independent tasks. Therefore we organize tasks of each trace file in the batch of 100. Last group may have less than 100 tasks. We apply each heuristic on each group consequently. Figure~\ref{fig:best_variants_batch} shows performance of the best variants of all categories for both applications. The plots exhibit behavior similar to Figures~\ref{fig:ratio_to_optimal_best_hf} and ~\ref{fig:ratio_to_optimal_best_ccsd}, static order with dynamic correction variants attain maximum communication-computation overlap and outperform other heuristics.
		%%		\todo[inline]{SKumar:combine both plots}
